Scalar on Function Regression

### 2022-01-04 Presentation/slides.tex
Added pca plots
Update slides.tex

	\begin{frame}{Introduction}
		Jona \\
		Introductory Example $\rightarrow$ Octane/NIR-spectrum
	\end{frame}

	\begin{frame}{Theory}
		Jona \\
		Motivation from multivariate regression (multivariate dgp).
	\end{frame}

@@

	\begin{frame}{Estimation via Basis Representation}
		Assume the following data generating process
		$$Y(\omega) = \alpha + \int_{0}^{1} \beta(s) F(\omega)(s) \mathrm{d}s + \epsilon(\omega)$$
		\begin{itemize}
			\item $Y$ and $\epsilon$ realize in $\mathbb{R}$ and $F$ realizes in $\mathbb{L}^2[0,1]$
		\end{itemize}
		\vspace{0.2cm}

		Assume that we have a data set containing observations each of which is made up of:
		\begin{itemize}
			\item $y_i$: \quad a scalar realizations of $Y$
			\item $f_i(t)$: \quad a realization of $F$
		\end{itemize}

	\begin{frame}{Estimation via Basis Representation}
		Let $\{b_i(t) \: \vert \: i = 1, \dots, \infty\}$ be a basis of $\mathbb{L}^2[0,1]$
		\vspace{0.2cm}

		Then we have the following representation of $\beta(t)$
		$$\beta(t) = \sum_{j = 1}^{\infty} \psi_j b_j(t) = \sum_{j = 1}^{L} \psi_j b_j(t) + \delta(t) \approx \sum_{j = 1}^{L} \psi_j b_j(t)$$
		and we can transform the data generating process into:
		\begin{equation}\notag
			\begin{split}
				Y(\omega) & = \alpha + \int_{0}^{1}\left[\left(\sum_{j = 1}^{\infty} \psi_j  b_j(s)\right) F(\omega)(s) \right]\mathrm{d}s + \epsilon(\omega) \\
						  & = \alpha + \sum_{j = 1}^{\infty} \left[\psi_j \textcolor{red}{\int_{0}^{1} F(\omega)(s) b_j(s)\mathrm{d}s}\right] + \epsilon(\omega)	  
			\end{split}
		\end{equation}
	\end{frame}

	\begin{frame}
		$$\int_{0}^{1} F(\omega)(s) b_j(s)\mathrm{d}s$$ 
		This is a scalar we can calculate for each observation and for each deterministic basis function $b_j(t)$.
	\end{frame}

	\begin{frame}{Theory - FPCA}
		Jakob
		\begin{itemize}
			\item Let's assume you know the theory of PCA (pc from varcov matrix) $\checkmark$
			\item Introduce mean and covariance functions of random functions $\checkmark$
			\item There is another cool basis $\rightarrow$ Eigenbasis (Karhunen-Loeve Expansion) $\checkmark$
			\item Sample Analog! (create a basis from observations and use for basis regression) $\checkmark$
			\item Plot fpcs and approximation of function realization
		\end{itemize}
	\end{frame}

@@

		The $\{\hat{\nu}_i(s) \: \vert \: i \in \mathcal{I}\}$ are called \textbf{Functional Principal Components} and can serve as a basis for representing the original curves. 
		\vspace{0.2cm}

		The corresponding scores $\hat{\xi}_i$ can be derived as
		$$\hat{\xi}_j(\omega) = \int_{0}^{1} (F(\omega)(s) - \hat{\mu}(s)) \hat{\nu}_j(s) \mathrm{d}s$$

@@

	\begin{frame}{Simulation Setup \& Application}
		Jona
		\begin{itemize}
			\item Compare b-spline / fourier regression chosen via criterion (cv/aic/...)
			\item Similar for fpca
			\item generate new curves from observed curves motivated by Karhunen-Loeve expansion
			\item Compare optimal variants with test and training sets
			\item Connect to Application
		\end{itemize}
	\end{frame}

>>>>

	\begin{frame}{Introduction}
		Jona \\	
		\begin{itemize}
			\item Near-infrared (NIR) spectroscopy enables fast diagnostics by using the NIR region of the electromagnetic spectrum
			\item Suited for field-monitoring / on-line analysis and diagnostics of e.g: prediction of octane ratings!

			\item Spectroscopy results in high-dimensional dataset.
			\item
			This set of measurements along a continuum can be viewed as set of smooth spectral curves
			\item
			Regression to determine relationship between octane rating and spectral curves
			\end{itemize}
	\end{frame}

	\begin{frame}{Theory}

		Motivation from multivariate regression (multivariate dgp).

		\begin{itemize}
			\item
			A simple functional dataset is given by $x_{n}(t_{j,n}) \in \mathbb{R}, t_{j,n} \in [T_1, T_2], n = 1,2,...,N, j = 1,2,..., J_n  $
			\item Continuous underlying process, where $x_n(t)$ exists at any point $t$, but is only observed at $x_{n}(t_{j,n})$
			\item NIR-spectroscopy measurements, financial data, medical data, human perception (musical pitch, ...)

			\item To abstract information from the curves, they must be interpretable!

			\end{itemize}
	\end{frame}

@@

	\begin{frame}{Theory - FPCA}
		Jakob
		\begin{itemize}
			\item Let's assume you know the theory of PCA (pc from varcov matrix)
			\item Introduce mean and covariance functions of random functions
			\item There is another cool basis $\rightarrow$ Eigenbasis (Karhunen-Loeve Expansion)
			\item Sample Analog! (create a basis from observations and use for basis regression)
			\item Plot fpcs and approximation of function realization
		\end{itemize}
	\end{frame}

@@

		The $\{\hat{\nu}_i(s) \: \vert \: i \in \mathcal{I}\}$ are then called the functional principal components and the corresponding scores $\xi_i$ can be derived using the following equation.
		$$\xi_j(\omega) = \int_{0}^{1} (F(\omega)(s) - \hat{\mu}(s)) \hat{\nu}_j(s) \mathrm{d}s$$

@@

	\begin{frame}{Simulation Setup \& Application}
		\begin{itemize}
		\item
		 Use the gasoline dataset (NIR-spectroscopy, 60 $\times$ 401) to predict octane ratings.
		\item 
		Generate new curves from gasoline dataset, motivated by Karhunen-Loeve expansion:

		$$F^{new}(\omega)(t) = \mu^{NIR}(t) + \sum_{j = 1}^{\infty} \xi_j^{new} \nu_j(t)^{new}$$ \\
		where \\
		 \vspace{0.1cm}
		$$\xi_{j}^{new} \in  \mathcal{N}_{n}(0,\Sigma)$$ \\
		and $\nu_j(t)^{new}$ be a $grid \times n_{FPC}$ matrix.

		\end{itemize}
	\end{frame}

	\begin{frame}{Simulation Setup \& Application cont.}
		\begin{itemize}
		\item
		Following Reiss and Ogden (2007), let $f_1$ and $f_2$ be two true coefficient functions that differ in smoothness: 
		\vspace{0.4cm}

   $f_1 = 2\sin(0.5\pi t) + 4\sin(1.5 \pi t) + 5\sin(2.5 \pi t)$\\
   \vspace{0.1cm}

    $	f_2 = 1.5^{\frac{-0,5(t-0.3)^2}{0.02^2}} - 4^{\frac{-0,5(t-0.45)^2}{0.015^2}} +  8^{\frac{-0,5(t-0.6)^2}{0.02^2}} -  1^{\frac{-0,5(t-0.8)^2}{0.03^2}}$

		\end{itemize}
		\vspace{0.1cm}
		\begin{figure}
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[height=3cm]{smooth_function.png}
  \captionof{figure}{$f_1$, smooth function}
  \label{fig:test1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[height=3cm]{bumpy_function.png}
  \captionof{figure}{$f_2$, bumpy function}
  \label{fig:test2}
\end{minipage}
\end{figure}

	\begin{frame}{Simulation Setup \& Application cont.}
		\begin{itemize}
		\item
    		Let \\
    		 \vspace{0.1cm}

    		$Y_{1,f} = NIR \times f + \mathcal{N}(0,1) \times var(NIR \times f)/0.9 - var(NIR \times f)$, \\
    		  $Y_{2,f} = NIR \times f + \mathcal{N}(0,1) \times var(NIR \times f)/0.9 - var(NIR \times f)$ be two responses for $f \in f_1, f_2$.		
    		 	\item
    		 	 Four combinations with different number of cubic basis-function $n_{basis} \in (5,6,...,25)$ to perform regression using basis expansion and the FPCR approach.
			\item Compare results via criteria (CV, Mallows CP,...)
			\item 
			{\color{green} add results here!}

		\end{itemize}
	\end{frame}

	\begin{frame}{Simulation Setup \& Application cont.}
		\begin{itemize}
		\item
    		Use insights from the simulation study to uncover dependence.
    		\item
    		Same setup, but without generated curves.
    		\item
    		Validation set approach:	Scores of testdata needs to be estimated by the trainingdata.  {\color{green} explain in detail?}
    		\item
    		Report results by MSE scaled by variance.
		\end{itemize}
	\end{frame}
